---
title: "100 Days of Code, Day 1"
author: "Subah Nuzhat"
date: "2023-07-20"
output: html_document
---

This is a R markdown file:


```{r}
#this is a code snippet
#comments like this are ignored
#insert code chunks with Ctrl+Alt+I
```
You can find a list of existing data sets in R for practice https://stat.ethz.ch/R-manual/R-devel/library/datasets/html/00Index.html 

<span style="color:red;">Good Luck!</span>

### Assignment
Everything that exists in R is an object. We create R objects by using the assignment operator: <-
```{r}
a <- 9 # the number 1 has been assigned to the variable a
b <- 8
c <- 17

```

### Arithmatic
```{r}

a + 3 #add
b - 2 #subtract
4*5 #multiply
100/c # divide
2**3 # raise to a power
a %% b	#Modular arithmetic - to find out modulus

```

### DATA TYPES:
* Numeric
* Integers
* Character
* Logical
* Complex
* Data structures
```{r}
class(1)
class("down")
```

### VECTORS 
Vectors are lists of values; to create a vector:
```{r}
codes <- c(380, 124, 818)
country <- c("italy", "canada", "egypt")

# named vectors
codes <- c(italy = 380, canada = 124, egypt = 818)
#OR
names(codes) <- country

```

Create index
```{r}
codes[2] #second index
codes[c(1,3)] #multiple index
codes[1:2] #multiple index in series
codes["canada"] #index with name
codes[c("egypt","italy")] #index with name, multiple

```

### Create Sequences
```{r}
1:80 # OR
seq(1, 80, 2)  
```
### Dataframes:
Tables are represented as data frames in R. \n
R has pre-built data sets:

```{r}
data("DNase")
```
To find the structure of the data
```{r}
str(DNase)
```
Show first six lines of the data set:
```{r}
head(DNase)
```
Find the table names
```{r}
names(DNase)
```

##### Accessor $
Access a specific column from the data
```{r}
concentration <- DNase$conc
#or
concentration <- DNase[['conc']]
length(concentration) # tells us how many objects are in the column
class(concentration)

DNase[1,3]#select particular value
DNase[1,] #selects entire row 1
DNase[,1] #selects column 1

### Selection using logical Operation ###

```


###SORTING
