---
title: "100 Days of Code Day 3"
author: "Subah Nuzhat"
date: "2023-07-26"
output: html_document
---
# DATA
Data set used today: https://www.kaggle.com/datasets/uom190346a/sleep-health-and-lifestyle-dataset

### R has pre-existing data sets
```{r}
#lists all the data sets present in R
data(package = .packages(all.available = TRUE))
```

### Let's load the sleep data (csv file)
```{r}

sleep_data <- read.csv(file ="Sleep_health_and_lifestyle_dataset.csv", header = TRUE,stringsAsFactors=FALSE )
#header = TRUE tells R that the first row contains headers
#stringsAsFactors turns strings into factors, the default is TRUE

#Structure of the data
str(sleep_data)

#short summary
head(sleep_data)
#tells us the dimension of the data
dim(sleep_data)

sleep_data[1:10, ] #first 10 lines

names(sleep_data) #column names

summary(sleep_data) #summarises information about the data
```

###### boxplots:
```{r}
steps <- sleep_data$Daily.Steps
boxplot(steps)

#average steps
mean(steps)
median(steps)
```
Let's look at the type of profession the individual's in the data have
```{r}
prof <- unique(sleep_data$Occupation)
prof
```
Let's look at the age of the individuals
```{r}
ages <- sleep_data$Age
unique(ages) 

### Take a look at the youngest participant

youngest <- which.min(ages) #index of youngest participant
youngest_age_1 <- sleep_data$Age[youngest] # returns age of youngest participant

youngest_age_2 <- min(ages) #youngest participant's age

youngest_age_1 == youngest_age_2

# we found two ways to look for the smallest value, the same can be done for the largest value (i.e. oldest participent)

oldest <- which.max(ages) #index of oldest participant
oldest_1 <- sleep_data$Age[oldest]
oldest_2 <- max(ages) #oldest participants
oldest_1 == oldest_2
```

### Sorting
lets sort the dataframe based on Sleep Duration
```{r}
sleep_duration_sorted <-sort(sleep_data$Sleep.Duration) #sorts values in the sleep duration column and returns it
sleep_duration_sorted #numeric


# the order function returns indexes of each row if the entire column was sorted (returns the index that sorts input vector)
sleep_duration_ordered <- order(sleep_data$Sleep.Duration)
sleep_duration_ordered

# Rank Function - For any given vector it returns a vector with the rank of the first entry, second entry, etc., of the input vector.

rank(sleep_data$Sleep.Duration)

```




let's select doctors under the age of 35
```{r}
doctor <- sleep_data$Occupation == "Doctor"
under_35 <- sleep_data$Age <= 35
ind <- which(doctor & under_35)
doctors_under_35 <- sleep_data[ind, ]
```

Let's subset the doctors_under_35 data set to have fewer columns so that it is easier to read, 

```{r}
doctors_subset <- doctors_under_35[, c(1, 2, 5, 6, 9:11 )]
names(doctors_subset) #the columns present in this data frame
```
Create separate columns for diastolic and systolic pressure
```{r}
doctors_subset <- mutate(doctors_subset, systolic = as.integer(substr(doctors_subset$Blood.Pressure, 1, 3)))

doctors_subset <- mutate(doctors_subset, diastolic = as.integer(substr(doctors_subset$Blood.Pressure, 5, 6)))

#removing the Blood.Pressure column
doctors_subset <- doctors_subset[-6] 

#confirm that all the new columns have been added
names(doctors_subset)

```

Let's see if any of the doctors under the age of 35 have high blood pressure

```{r}
# But first let's look at the different values in the table
unique(doctors_subset$systolic)
# so apparently they either have a systolic blood pressure of 120 or 125

```
Does anyone have low diastolic pressure?
```{r}
unique(doctors_subset$diastolic) 
# 80
```
